### 2015-07-01 paper/sec-appendix-overlap.tex
Work on paper

\section{On the role of the overlap between different categories}
\label{sec:overlap}

In this appendix we discuss the role of the overlap
between  different categories.