### 2015-07-13 paper/HH4b.tex
Updated common plot style

\documentclass[a4paper,11pt]{article}
\usepackage{jheppub}
\usepackage{graphicx}% Include figure files
\usepackage{graphics}
\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}% bold math
\usepackage{epstopdf}
\usepackage{mathrsfs}
\usepackage{amssymb}
\usepackage{amsmath}
%\usepackage[pdftex]{hyperref}
\usepackage{natbib}

\def\bs{\boldsymbol}
\def\del{\partial}

%\def\p{{\boldsymbol p}_{\perp }}
%\def\pb{\bar {\boldsymbol p}_{\perp }}
%\def\pp{{\boldsymbol p}_{\perp }}
\def\p{{\boldsymbol p}}
\def\pb{\bar {\boldsymbol p}}
\def\pp{{\boldsymbol p}}
%\def\q{{\boldsymbol q}_\perp}
\def\q{{\boldsymbol q}}
\def\l{{\boldsymbol l}_\perp}
%\def\k{{\boldsymbol k}_\perp}
\def\k{{\boldsymbol k}}
\def\m{{\boldsymbol m}_\perp}
%\def\x{{\boldsymbol x}_\perp}
\def\x{{\boldsymbol x}}
\def\y{{\boldsymbol y}_\perp}
\def\X{{\boldsymbol X}_\perp}
\def\Y{{\boldsymbol Y}_\perp}
\def\D{{\boldsymbol D}_\perp}
\def\r{{\boldsymbol r}_\perp}
\def\z{{\boldsymbol z}_\perp}
%\def\v{{\boldsymbol v}_\perp}
\def\v{{\boldsymbol v}}
\def\w{{\boldsymbol w}_\perp}
\def\b{{\boldsymbol b}_\perp}
\def\Q{{\boldsymbol Q}_\perp}
\def\M{{\boldsymbol M}_\perp}
%\def\bkappa{{\boldsymbol \kappa}_\perp}
%\def\bbkappa{\bar{\boldsymbol \kappa}_\perp}
\def\bkappa{{\boldsymbol \kappa}}
\def\bbkappa{\bar{\boldsymbol \kappa}}
\def\bnu{{\boldsymbol \nu}_\perp}
\def\V{\hat{\boldsymbol v}_{1\perp}}
\def\K{\hat{\boldsymbol k}_{1\perp}}
\def\bV{\hat{\boldsymbol v}_{2\perp}}
\def\bK{\hat{\boldsymbol k}_{2\perp}}
\def\qqb{{q\bar q}}
\def\sM{{\scriptscriptstyle M}}

\newcommand{\beq}{\begin{eqnarray}}
\newcommand{\eeq}{\end{eqnarray}}
\newcommand{\be}{\begin{equation}}
\newcommand{\ee}{\end{equation}}
\newcommand{\nn}{\nonumber\\ }
\newcommand{\labe}{\label}
\newcommand{\bea}{\begin{eqnarray}}
\newcommand{\eea}{\end{eqnarray}}

\newcommand{\la}{\left\langle}
\newcommand{\ra}{\right\rangle}
\newcommand{\lc}{\left[}
\newcommand{\rc}{\right]}
\newcommand{\lp}{\left(}
\newcommand{\rp}{\right)}

\bibliographystyle{JHEP}
\usepackage{multirow}

\title{Boosting Higgs Pair Production in the $b\bar{b}b\bar{b}$ Channel at the LHC with
  Multivariate Techniques}

\author{Katharina Behr, Nurfiki Bin Norjoharuddeen, Daniela Bortoletto, James Frost, Nathan P. Hartland, Cigdem Issever and Juan Rojo}

\affiliation{Physics Department, 1 Keble Road, University of Oxford, United Kingdom }

\preprint{OUTP-15-14P}

\abstract{
  The measurement of Higgs pair production is one of the cornerstones of
  the LHC program for the next years.
  %
  Double Higgs production provides a unique window to the electroweak
  symmetry breaking mechanism, and in particular it allows
  to access the Higgs trilinear coupling.
  %
  In this work we study the feasibility of the observation of the production of
  Higgs boson pairs at the High-Luminosity LHC (HL-LHC) in the
  $b\bar{b}b\bar{b}$ final state.
  %
  While the signal in this channel
  is enhanced by the large branching fraction of Higgs bosons to
  $b\bar{b}$ pairs, the measurement is  challenging due to the
  overwhelming QCD multijet background.
  %
  Our analysis is based on a combination of traditional cut-based
  methods with modern multivariate techniques, which lead to a substantial
  enhacement of the signal discrimination.
  %
  We account for  all relevant
  backgrounds, including the contribution from light jet fakes that is found
  to be as important or even dominant over the irreducible $4b$ QCD background.
  %
  By combining the contribution from the boosted, intermediate and resolved
  topologies we achieve a signal significance of $S/\sqrt{B}\simeq 3.2$.
  %
  We therefore demonstrate that at the HL-LHC the $b\bar{b}b\bar{b}$ channel alone
  should allow to claim evidence of the observation of double Higgs production.
}

\begin{document} 
\maketitle

\flushbottom

\input{sec-intro}
\input{sec-mcsamples}
\input{sec-analysis}
\input{sec-results}
\input{sec-mva}
\input{sec-optimisation}
\input{sec-conclusions}

\bibliography{HH4b}

\end{document}