### 2016-01-21 paper-arxiv-v2/HH4b.tex
Revised version of paper

\documentclass[a4paper,11pt]{article}
\pdfoutput=1
%\usepackage{jheppub}
\usepackage{graphicx}% Include figure files
\usepackage{graphics}
\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}% bold math
\usepackage{epstopdf}
\usepackage{mathrsfs}
\usepackage{amssymb}
\usepackage{amsmath}
%\usepackage[pdftex]{hyperref}
%\usepackage{natbib}
\usepackage{multirow,cite}

\def\bs{\boldsymbol}
\def\del{\partial}

%\def\p{{\boldsymbol p}_{\perp }}
%\def\pb{\bar {\boldsymbol p}_{\perp }}
%\def\pp{{\boldsymbol p}_{\perp }}
\def\p{{\boldsymbol p}}
\def\pb{\bar {\boldsymbol p}}
\def\pp{{\boldsymbol p}}
%\def\q{{\boldsymbol q}_\perp}
\def\q{{\boldsymbol q}}
\def\l{{\boldsymbol l}_\perp}
%\def\k{{\boldsymbol k}_\perp}
\def\k{{\boldsymbol k}}
\def\m{{\boldsymbol m}_\perp}
%\def\x{{\boldsymbol x}_\perp}
\def\x{{\boldsymbol x}}
\def\y{{\boldsymbol y}_\perp}
\def\X{{\boldsymbol X}_\perp}
\def\Y{{\boldsymbol Y}_\perp}
\def\D{{\boldsymbol D}_\perp}
\def\r{{\boldsymbol r}_\perp}
\def\z{{\boldsymbol z}_\perp}
%\def\v{{\boldsymbol v}_\perp}
\def\v{{\boldsymbol v}}
\def\w{{\boldsymbol w}_\perp}
\def\b{{\boldsymbol b}_\perp}
\def\Q{{\boldsymbol Q}_\perp}
\def\M{{\boldsymbol M}_\perp}
%\def\bkappa{{\boldsymbol \kappa}_\perp}
%\def\bbkappa{\bar{\boldsymbol \kappa}_\perp}
\def\bkappa{{\boldsymbol \kappa}}
\def\bbkappa{\bar{\boldsymbol \kappa}}
\def\bnu{{\boldsymbol \nu}_\perp}
\def\V{\hat{\boldsymbol v}_{1\perp}}
\def\K{\hat{\boldsymbol k}_{1\perp}}
\def\bV{\hat{\boldsymbol v}_{2\perp}}
\def\bK{\hat{\boldsymbol k}_{2\perp}}
\def\qqb{{q\bar q}}
\def\sM{{\scriptscriptstyle M}}

\newcommand{\beq}{\begin{eqnarray}}
\newcommand{\eeq}{\end{eqnarray}}
\newcommand{\be}{\begin{equation}}
\newcommand{\ee}{\end{equation}}
\newcommand{\nn}{\nonumber\\ }
\newcommand{\labe}{\label}
\newcommand{\bea}{\begin{eqnarray}}
\newcommand{\eea}{\end{eqnarray}}

\newcommand{\la}{\left\langle}
\newcommand{\ra}{\right\rangle}
\newcommand{\lc}{\left[}
\newcommand{\rc}{\right]}
\newcommand{\lp}{\left(}
\newcommand{\rp}{\right)}

%\bibliographystyle{JHEP}

%\textwidth=15.0cm \textheight=23.5cm
\textwidth=16.0cm \textheight=23.0cm 
\topmargin 0cm \oddsidemargin 0cm 
\setlength{\unitlength}{1mm}

\usepackage{url}
\usepackage{hyperref}

\begin{document}

\vspace{-2.0cm}
\begin{flushright}
OUTP-15-14P
\end{flushright}

\begin{center}
  {\Large \bf Boosting Higgs Pair Production in the $b\bar{b}b\bar{b}$ Final State\\[0.2cm]
    with
  Multivariate Techniques}
\vspace{.7cm}

J. Katharina Behr, Daniela Bortoletto, James A. Frost,
  Nathan P. Hartland, \\Cigdem Issever and Juan Rojo

\vspace{.3cm}
{\it Physics Department, 1 Keble Road, University of Oxford, United Kingdom }

\vspace{1cm}

{\bf \large Abstract}\\
\end{center}
\vspace{0.1cm}
The measurement of Higgs pair production will be a cornerstone of
  the LHC program in the coming years.
  %
  Double Higgs production provides a crucial window upon the mechanism of electroweak
  symmetry breaking and has a unique sensitivity
  to the Higgs trilinear coupling.
  %
  We study the feasibility of a measurement of
  Higgs pair production in the
  $b\bar{b}b\bar{b}$ final state at the LHC.
  %
  Our analysis is based on a combination of traditional cut-based
  methods with state-of-the-art
  multivariate techniques.
  %
  We account for all relevant
  backgrounds, including the contributions from light and charm
  jet mis-identification, which are ultimately comparable in size 
  to the irreducible $4b$ QCD background.
  %
  We demonstrate the robustness of our analysis strategy in
  a high pileup environment.
  %
  For an integrated luminosity of $\mathcal{L}=3$ ab$^{-1}$,
  a signal significance of  $S/\sqrt{B}\simeq 3$ is obtained,
  indicating
  that the $b\bar{b}b\bar{b}$ final state
alone could allow for the observation of double Higgs production
at the High Luminosity LHC.

\clearpage

\input{sec-intro}
\input{sec-mcsamples}
\input{sec-analysis}
\input{sec-results}
\input{sec-mva}
\input{sec-conclusions}

%%%%%%%%%%%%%%%%%%

\providecommand{\href}[2]{#2}\begingroup\raggedright\begin{thebibliography}{100}

\bibitem{ATLAS:2013hta}
{\bf ATLAS} Collaboration, {\it {Physics at a High-Luminosity LHC with ATLAS}},
   \href{http://arxiv.org/abs/1307.7292}{{\tt arXiv:1307.7292}}.

\bibitem{CMS:2013xfa}
{\bf CMS} Collaboration, {\it {Projected Performance of an Upgraded CMS
  Detector at the LHC and HL-LHC: Contribution to the Snowmass Process}},
  \href{http://arxiv.org/abs/1307.7135}{{\tt arXiv:1307.7135}}.

\bibitem{baglio}
J.~Baglio, A.~Djouadi, R.~Grober, M.~Muhlleitner, J.~Quevillon, et~al., {\it
  {The measurement of the Higgs self-coupling at the LHC: theoretical status}},
   {\em JHEP} {\bf 1304} (2013) 151,
  [\href{http://arxiv.org/abs/1212.5581}{{\tt arXiv:1212.5581}}].

\bibitem{Giudice:2007fh}
G.~Giudice, C.~Grojean, A.~Pomarol, and R.~Rattazzi, {\it {The
  Strongly-Interacting Light Higgs}},  {\em JHEP} {\bf 0706} (2007) 045,
  [\href{http://arxiv.org/abs/hep-ph/0703164}{{\tt hep-ph/0703164}}].

\bibitem{Contino:2010mh}
R.~Contino, C.~Grojean, M.~Moretti, F.~Piccinini, and R.~Rattazzi, {\it {Strong
  Double Higgs Production at the LHC}},  {\em JHEP} {\bf 1005} (2010) 089,
  [\href{http://arxiv.org/abs/1002.1011}{{\tt arXiv:1002.1011}}].

\bibitem{Aad:2015xja}
{\bf ATLAS} Collaboration, G.~Aad et~al., {\it {Searches for Higgs boson pair
  production in the $hh\to bb\tau\tau, \gamma\gamma WW^*, \gamma\gamma bb,
  bbbb$ channels with the ATLAS detector}},  {\em Phys. Rev.} {\bf D92} (2015)
  092004, [\href{http://arxiv.org/abs/1509.04670}{{\tt arXiv:1509.04670}}].

\bibitem{Aad:2015uka}
{\bf ATLAS} Collaboration, G.~Aad et~al., {\it {Search for Higgs boson pair
  production in the $b\bar{b}b\bar{b}$ final state from pp collisions at
  $\sqrt{s} = 8$ TeVwith the ATLAS detector}},  {\em Eur. Phys. J.} {\bf C75}
  (2015), no.~9 412, [\href{http://arxiv.org/abs/1506.00285}{{\tt
  arXiv:1506.00285}}].

\bibitem{Aad:2014yja}
{\bf ATLAS} Collaboration, G.~Aad et~al., {\it {Search For Higgs Boson Pair
  Production in the $\gamma\gamma b\bar{b}$ Final State using $pp$ Collision
  Data at $\sqrt{s}=8$ TeV from the ATLAS Detector}},  {\em Phys.Rev.Lett.}
  {\bf 114} (2015), no.~8 081802, [\href{http://arxiv.org/abs/1406.5053}{{\tt
  arXiv:1406.5053}}].

\bibitem{Khachatryan:2015yea}
{\bf CMS} Collaboration, V.~Khachatryan et~al., {\it {Search for resonant pair
  production of Higgs bosons decaying to two bottom quark–antiquark pairs in
  proton–proton collisions at 8 TeV}},  {\em Phys. Lett.} {\bf B749} (2015)
  560--582, [\href{http://arxiv.org/abs/1503.04114}{{\tt arXiv:1503.04114}}].

\bibitem{Chatrchyan:2011wt}
{\bf CMS} Collaboration, S.~Chatrchyan et~al., {\it {Search for resonant HH
  production in $2\gamma2b$ channel}}, {\tt CMS-PAS-HIG-13-032}.

\bibitem{Arkani-Hamed:2015vfh}
N.~Arkani-Hamed, T.~Han, M.~Mangano, and L.-T. Wang, {\it {Physics
  Opportunities of a 100 TeV Proton-Proton Collider}},
  \href{http://arxiv.org/abs/1511.06495}{{\tt arXiv:1511.06495}}.

\bibitem{Barr:2014sga}
A.~J. Barr, M.~J. Dolan, C.~Englert, D.~E. Ferreira~de Lima, and M.~Spannowsky,
  {\it {Higgs Self-Coupling Measurements at a 100 TeV Hadron Collider}},  {\em
  JHEP} {\bf 1502} (2015) 016, [\href{http://arxiv.org/abs/1412.7154}{{\tt
  arXiv:1412.7154}}].

\bibitem{Papaefstathiou:2015iba}
A.~Papaefstathiou, {\it {Discovering Higgs boson pair production through rare
  final states at a 100 TeV collider}},  {\em Phys.Rev.} {\bf D91} (2015),
  no.~11 113016, [\href{http://arxiv.org/abs/1504.04621}{{\tt
  arXiv:1504.04621}}].

\bibitem{Azatov:2015oxa}
A.~Azatov, R.~Contino, G.~Panico, and M.~Son, {\it {Effective field theory
  analysis of double Higgs boson production via gluon fusion}},  {\em Phys.
  Rev.} {\bf D92} (2015), no.~3 035001,
  [\href{http://arxiv.org/abs/1502.00539}{{\tt arXiv:1502.00539}}].

\bibitem{Contino:2013gna}
R.~Contino, C.~Grojean, D.~Pappadopulo, R.~Rattazzi, and A.~Thamm, {\it {Strong
  Higgs Interactions at a Linear Collider}},  {\em JHEP} {\bf 1402} (2014) 006,
  [\href{http://arxiv.org/abs/1309.7038}{{\tt arXiv:1309.7038}}].

\bibitem{Dittmaier:2012vm}
{LHC Higgs Cross Section Working Group}, S.~Dittmaier, C.~Mariotti,
  G.~Passarino, and R.~Tanaka~(Eds.), {\it {Handbook of LHC Higgs Cross
  Sections: 2. Differential Distributions}},  {\em CERN-2012-002} (CERN,
  Geneva, 2012) [\href{http://arxiv.org/abs/1201.3084}{{\tt arXiv:1201.3084}}].

\bibitem{Frederix:2014hta}
R.~Frederix, S.~Frixione, V.~Hirschi, F.~Maltoni, O.~Mattelaer, P.~Torrielli,
  E.~Vryonidou, and M.~Zaro, {\it {Higgs pair production at the LHC with NLO
  and parton-shower effects}},  {\em Phys. Lett.} {\bf B732} (2014) 142--149,
  [\href{http://arxiv.org/abs/1401.7340}{{\tt arXiv:1401.7340}}].

\bibitem{deFlorian:2013jea}
D.~de~Florian and J.~Mazzitelli, {\it {Higgs Boson Pair Production at
  Next-to-Next-to-Leading Order in QCD}},  {\em Phys. Rev. Lett. 111,} {\bf
  201801} (2013) 201801, [\href{http://arxiv.org/abs/1309.6594}{{\tt
  arXiv:1309.6594}}].

\bibitem{deFlorian:2015moa}
D.~de~Florian and J.~Mazzitelli, {\it {Higgs pair production at
  next-to-next-to-leading logarithmic accuracy at the LHC}},  {\em JHEP} {\bf
  09} (2015) 053, [\href{http://arxiv.org/abs/1505.07122}{{\tt
  arXiv:1505.07122}}].

\bibitem{Baur:2003gp}
U.~Baur, T.~Plehn, and D.~L. Rainwater, {\it {Probing the Higgs selfcoupling at
  hadron colliders using rare decays}},  {\em Phys.Rev.} {\bf D69} (2004)
  053004, [\href{http://arxiv.org/abs/hep-ph/0310056}{{\tt hep-ph/0310056}}].

\bibitem{Barger:2013jfa}
V.~Barger, L.~L. Everett, C.~Jackson, and G.~Shaughnessy, {\it {Higgs-Pair
  Production and Measurement of the Triscalar Coupling at LHC(8,14)}},
  \href{http://arxiv.org/abs/1311.2931}{{\tt arXiv:1311.2931}}.

\bibitem{Lu:2015jza} 
  C.~T.~Lu, J.~Chang, K.~Cheung and J.~S.~Lee,
  {\it An exploratory study of Higgs-boson pair production},
  JHEP {\bf 1508}, 133 (2015),
  \href{http://arxiv.org/abs/1505.00957}{{\tt arXiv:1505.00957}}.

\bibitem{Baur:2003gpa}
U.~Baur, T.~Plehn, and D.~L. Rainwater, {\it {Examining the Higgs boson
  potential at lepton and hadron colliders: A Comparative analysis}},  {\em
  Phys.Rev.} {\bf D68} (2003) 033001,
  [\href{http://arxiv.org/abs/hep-ph/0304015}{{\tt hep-ph/0304015}}].

\bibitem{Barr:2013tda}
A.~J. Barr, M.~J. Dolan, C.~Englert, and M.~Spannowsky, {\it {Di-Higgs final
  states augMT2ed -- selecting $hh$ events at the high luminosity LHC}},  {\em
  Phys. Lett.} {\bf B728} (2014) 308--313,
  [\href{http://arxiv.org/abs/1309.6318}{{\tt arXiv:1309.6318}}].

\bibitem{Dolan:2012rv}
M.~J. Dolan, C.~Englert, and M.~Spannowsky, {\it {Higgs self-coupling
  measurements at the LHC}},  {\em JHEP} {\bf 1210} (2012) 112,
  [\href{http://arxiv.org/abs/1206.5001}{{\tt arXiv:1206.5001}}].

\bibitem{Dolan:2013rja}
M.~J. Dolan, C.~Englert, N.~Greiner, and M.~Spannowsky, {\it {Further on up the
  road: $hhjj$ production at the LHC}},  {\em Phys.Rev.Lett.} {\bf 112} (2014)
  101802, [\href{http://arxiv.org/abs/1310.1084}{{\tt arXiv:1310.1084}}].

\bibitem{Papaefstathiou:2012qe}
A.~Papaefstathiou, L.~L. Yang, and J.~Zurita, {\it {Higgs boson pair production
  at the LHC in the $b \bar{b} W^+ W^-$ channel}},  {\em Phys.Rev.} {\bf D87}
  (2013) 011301, [\href{http://arxiv.org/abs/1209.1489}{{\tt
  arXiv:1209.1489}}].

\bibitem{Wardrope:2014kya}
D.~Wardrope, E.~Jansen, N.~Konstantinidis, B.~Cooper, R.~Falla, et~al., {\it
  {Non-resonant Higgs-pair production in the $b\overline{b}$ $b\overline{b}$
  final state at the LHC}},  {\em Eur.Phys.J.} {\bf C75} (2015), no.~5 219,
  [\href{http://arxiv.org/abs/1410.2794}{{\tt arXiv:1410.2794}}].

\bibitem{deLima:2014dta}
D.~E. Ferreira~de Lima, A.~Papaefstathiou, and M.~Spannowsky, {\it {Standard
  model Higgs boson pair production in the ( $ b\overline{b} $ )( $
  b\overline{b} $ ) final state}},  {\em JHEP} {\bf 1408} (2014) 030,
  [\href{http://arxiv.org/abs/1404.7139}{{\tt arXiv:1404.7139}}].

\bibitem{Dolan:2015zja}
M.~J. Dolan, C.~Englert, N.~Greiner, K.~Nordstrom, and M.~Spannowsky, {\it
  {$hhjj$ production at the LHC}},  {\em Eur. Phys. J.} {\bf C75} (2015), no.~8
  387, [\href{http://arxiv.org/abs/1506.08008}{{\tt arXiv:1506.08008}}].

\bibitem{Brooijmans:2014eja}
G.~Brooijmans, R.~Contino, B.~Fuks, F.~Moortgat, P.~Richardson, et~al., {\it
  {Les Houches 2013: Physics at TeV Colliders: New Physics Working Group
  Report}},  \href{http://arxiv.org/abs/1405.1617}{{\tt arXiv:1405.1617}}.

\bibitem{Barger:1988jk}
V.~D. Barger, T.~Han, and R.~J.~N. Phillips, {\it {Double Higgs Boson
  Bremsstrahlung From $W$ and $Z$ Bosons at Supercolliders}},  {\em Phys. Rev.}
  {\bf D38} (1988) 2766.

\bibitem{Cao:2015oxx}
Q.-H. Cao, Y.~Liu, and B.~Yan, {\it {Measuring Trilinear Higgs Coupling in
  $WHH$ and $ZHH$ Productions at the HL-LHC}},
  \href{http://arxiv.org/abs/1511.03311}{{\tt arXiv:1511.03311}}.

\bibitem{Englert:2014uqa}
C.~Englert, F.~Krauss, M.~Spannowsky, and J.~Thompson, {\it {Di-Higgs
  phenomenology in $t\bar{t}hh$: The forgotten channel}},  {\em Phys.Lett.}
  {\bf B743} (2015) 93--97, [\href{http://arxiv.org/abs/1409.8074}{{\tt
  arXiv:1409.8074}}].

\bibitem{Liu-Sheng:2014gxa}
L.-S. Ling, R.-Y. Zhang, W.-G. Ma, L.~Guo, W.-H. Li, et~al., {\it {NNLO QCD
  corrections to Higgs pair production via vector boson fusion at hadron
  colliders}},  {\em Phys.Rev.} {\bf D89} (2014), no.~7 073001,
  [\href{http://arxiv.org/abs/1401.7754}{{\tt arXiv:1401.7754}}].

\bibitem{Slawinska:2014vpa}
M.~Slawinska, W.~van~den Wollenberg, B.~van Eijk, and S.~Bentvelsen, {\it
  {Phenomenology of the trilinear Higgs coupling at proton-proton colliders}},
  \href{http://arxiv.org/abs/1408.5010}{{\tt arXiv:1408.5010}}.

\bibitem{Chen:2014xra}
C.-R. Chen and I.~Low, {\it {Double take on new physics in double Higgs boson
  production}},  {\em Phys.Rev.} {\bf D90} (2014), no.~1 013018,
  [\href{http://arxiv.org/abs/1405.7040}{{\tt arXiv:1405.7040}}].

\bibitem{Goertz:2013kp}
F.~Goertz, A.~Papaefstathiou, L.~L. Yang, and J.~Zurita, {\it {Higgs Boson
  self-coupling measurements using ratios of cross sections}},  {\em JHEP} {\bf
  1306} (2013) 016, [\href{http://arxiv.org/abs/1301.3492}{{\tt
  arXiv:1301.3492}}].

\bibitem{Dawson:2015oha}
S.~Dawson, A.~Ismail, and I.~Low, {\it {What’s in the loop? The anatomy of
  double Higgs production}},  {\em Phys.Rev.} {\bf D91} (2015), no.~11 115008,
  [\href{http://arxiv.org/abs/1504.05596}{{\tt arXiv:1504.05596}}].

\bibitem{Maltoni:2014eza}
F.~Maltoni, E.~Vryonidou, and M.~Zaro, {\it {Top-quark mass effects in double
  and triple Higgs production in gluon-gluon fusion at NLO}},  {\em JHEP} {\bf
  1411} (2014) 079, [\href{http://arxiv.org/abs/1408.6542}{{\tt
  arXiv:1408.6542}}].

\bibitem{Maierhofer:2013sha}
P.~Maierhöfer and A.~Papaefstathiou, {\it {Higgs Boson pair production merged
  to one jet}},  {\em JHEP} {\bf 1403} (2014) 126,
  [\href{http://arxiv.org/abs/1401.0007}{{\tt arXiv:1401.0007}}].

\bibitem{Grigo:2013rya}
J.~Grigo, J.~Hoff, K.~Melnikov, and M.~Steinhauser, {\it {On the Higgs boson
  pair production at the LHC}},  {\em Nucl.Phys.} {\bf B875} (2013) 1--17,
  [\href{http://arxiv.org/abs/1305.7340}{{\tt arXiv:1305.7340}}].

\bibitem{Grigo:2014jma}
J.~Grigo, K.~Melnikov, and M.~Steinhauser, {\it {Virtual corrections to Higgs
  boson pair production in the large top quark mass limit}},  {\em Nucl. Phys.}
  {\bf B888} (2014) 17--29, [\href{http://arxiv.org/abs/1408.2422}{{\tt
  arXiv:1408.2422}}].

\bibitem{Nishiwaki:2013cma}
K.~Nishiwaki, S.~Niyogi, and A.~Shivaji, {\it {$ttH$ Anomalous Coupling in
  Double Higgs Production}},  {\em JHEP} {\bf 1404} (2014) 011,
  [\href{http://arxiv.org/abs/1309.6907}{{\tt arXiv:1309.6907}}].

\bibitem{Cao:2015oaa} 
  Q.~H.~Cao, B.~Yan, D.~M.~Zhang and H.~Zhang, 
  {\it Resolving the Degeneracy in Single Higgs Production with Higgs Pair Production}, 
  {\em Phys.\ Lett.\ B} {\bf 752}, 285 (2016),
  [\href{http://arxiv.org/abs/1508.06512}{{\tt arXiv:1508.06512}}].

\bibitem{Dall'Osso:2015aia}
M.~Dall'Osso, T.~Dorigo, C.~A. Gottardo, A.~Oliveira, M.~Tosi, et~al., {\it
  {Higgs Pair Production: Choosing Benchmarks With Cluster Analysis}},
  \href{http://arxiv.org/abs/1507.02245}{{\tt arXiv:1507.02245}}.

\bibitem{Liu:2014rba}
N.~Liu, S.~Hu, B.~Yang, and J.~Han, {\it {Impact of top-Higgs couplings on
  Di-Higgs production at future colliders}},  {\em JHEP} {\bf 1501} (2015) 008,
  [\href{http://arxiv.org/abs/1408.4191}{{\tt arXiv:1408.4191}}].

\bibitem{Goertz:2014qta}
F.~Goertz, A.~Papaefstathiou, L.~L. Yang, and J.~Zurita, {\it {Higgs boson pair
  production in the D=6 extension of the SM}},  {\em JHEP} {\bf 1504} (2015)
  167, [\href{http://arxiv.org/abs/1410.3471}{{\tt arXiv:1410.3471}}].

\bibitem{He:2015spf}
H.-J. He, J.~Ren, and W.~Yao, {\it {Probing New Physics of Cubic Higgs
  Interaction via Higgs Pair Production at Hadron Colliders}},
  \href{http://arxiv.org/abs/1506.03302}{{\tt arXiv:1506.03302}}.

\bibitem{Grober:2015cwa}
R.~Grober, M.~Muhlleitner, M.~Spira, and J.~Streicher, {\it {NLO QCD
  Corrections to Higgs Pair Production including Dimension-6 Operators}},  {\em
  JHEP} {\bf 09} (2015) 092, [\href{http://arxiv.org/abs/1504.06577}{{\tt
  arXiv:1504.06577}}].

\bibitem{Gouzevitch:2013qca}
M.~Gouzevitch, A.~Oliveira, J.~Rojo, R.~Rosenfeld, G.~P. Salam, et~al., {\it
  {Scale-invariant resonance tagging in multijet events and new physics in
  Higgs pair production}},  {\em JHEP} {\bf 1307} (2013) 148,
  [\href{http://arxiv.org/abs/1303.6636}{{\tt arXiv:1303.6636}}].

\bibitem{Cooper:2013kia}
B.~Cooper, N.~Konstantinidis, L.~Lambourne, and D.~Wardrope, {\it {Boosted $hh
  → b\overline{b}b\overline{b}$: A new topology in searches for TeV-scale
  resonances at the LHC}},  {\em Phys. Rev.} {\bf D88} (2013), no.~11 114005,
  [\href{http://arxiv.org/abs/1307.0407}{{\tt arXiv:1307.0407}}].

\bibitem{No:2013wsa}
J.~M. No and M.~Ramsey-Musolf, {\it {Probing the Higgs Portal at the LHC
  Through Resonant di-Higgs Production}},  {\em Phys.Rev.} {\bf D89} (2014),
  no.~9 095031, [\href{http://arxiv.org/abs/1310.6035}{{\tt arXiv:1310.6035}}].

\bibitem{Wen-Juan:2015gqg}
Z.~Wen-Juan, M.~Wen-Gan, Z.~Ren-You, L.~Xiao-Zhou, G.~Lei, and C.~Chong, {\it
  {Double Higgs boson production and decay in Randall-Sundrum model at hadron
  colliders}},  \href{http://arxiv.org/abs/1512.01766}{{\tt arXiv:1512.01766}}.

\bibitem{Belyaev:1999kk}
A.~Belyaev, M.~Drees, O.~J. Eboli, J.~Mizukoshi, and S.~Novaes, {\it
  {Supersymmetric Higgs pair discovery prospects at hadron colliders}},
  \href{http://arxiv.org/abs/hep-ph/9910400}{{\tt hep-ph/9910400}}.

\bibitem{Han:2013sga}
C.~Han, X.~Ji, L.~Wu, P.~Wu, and J.~M. Yang, {\it {Higgs pair production with
  SUSY QCD correction: revisited under current experimental constraints}},
  {\em JHEP} {\bf 1404} (2014) 003, [\href{http://arxiv.org/abs/1307.3790}{{\tt
  arXiv:1307.3790}}].

\bibitem{Hespel:2014sla}
B.~Hespel, D.~Lopez-Val, and E.~Vryonidou, {\it {Higgs pair production via
  gluon fusion in the Two-Higgs-Doublet Model}},  {\em JHEP} {\bf 1409} (2014)
  124, [\href{http://arxiv.org/abs/1407.0281}{{\tt arXiv:1407.0281}}].

\bibitem{Wu:2015nba}
L.~Wu, J.~M. Yang, C.-P. Yuan, and M.~Zhang, {\it {Higgs self-coupling in the
  MSSM and NMSSM after the LHC Run 1}},  {\em Phys.Lett.} {\bf B747} (2015)
  378--389, [\href{http://arxiv.org/abs/1504.06932}{{\tt arXiv:1504.06932}}].

\bibitem{Cao:2014kya}
J.~Cao, D.~Li, L.~Shang, P.~Wu, and Y.~Zhang, {\it Exploring the Higgs Sector
  of a Most Natural NMSSM and its Prediction on Higgs Pair Production at the
  LHC},  {\em JHEP} {\bf 1412} (2014) 026,
  [\href{http://arxiv.org/abs/1409.8431}{{\tt arXiv:1409.8431}}].

\bibitem{Ellwanger:2013ova}
U.~Ellwanger, {\it {Higgs pair production in the NMSSM at the LHC}},  {\em
  JHEP} {\bf 1308} (2013) 077, [\href{http://arxiv.org/abs/1306.5541}{{\tt
  arXiv:1306.5541}}].

%\cite{Cao:2013si}
\bibitem{Cao:2013si} 
  J.~Cao, Z.~Heng, L.~Shang, P.~Wan and J.~M.~Yang,
  {\it Pair Production of a 125 GeV Higgs Boson in MSSM and NMSSM at the LHC,}
  {\em JHEP} {\bf 1304}, 134 (2013),
  [\href{http://arxiv.org/abs/1301.6437}{{\tt
  arXiv:1301.6437}}].

\bibitem{Khachatryan:2015tha}
{\bf CMS} Collaboration, V.~Khachatryan et~al., {\it {Searches for a heavy
  scalar boson $ \mathrm{H} $ decaying to a pair of 125 GeV Higgs bosons $
  \mathrm{ hh } $ or for a heavy pseudoscalar boson $ \mathrm{A} $ decaying to
  $ \mathrm{Zh} $, in the final states with $\mathrm{h} \to \tau \tau$}},
  \href{http://arxiv.org/abs/1510.01181}{{\tt arXiv:1510.01181}}.

\bibitem{Alwall:2014hca}
J.~Alwall, R.~Frederix, S.~Frixione, V.~Hirschi, F.~Maltoni, et~al., {\it {The
  automated computation of tree-level and next-to-leading order differential
  cross sections, and their matching to parton shower simulations}},  {\em
  JHEP} {\bf 1407} (2014) 079, [\href{http://arxiv.org/abs/1405.0301}{{\tt
  arXiv:1405.0301}}].

\bibitem{Plehn:1996wb}
T.~Plehn, M.~Spira, and P.~Zerwas, {\it {Pair production of neutral Higgs
  particles in gluon-gluon collisions}},  {\em Nucl.Phys.} {\bf B479} (1996)
  46--64, [\href{http://arxiv.org/abs/hep-ph/9603205}{{\tt hep-ph/9603205}}].

\bibitem{Hirschi:2015iia}
V.~Hirschi and O.~Mattelaer, {\it {Automated event generation for loop-induced
  processes}},  {\em JHEP} {\bf 10} (2015) 146,
  [\href{http://arxiv.org/abs/1507.00020}{{\tt arXiv:1507.00020}}].

\bibitem{Ball:2014uwa}
{\bf NNPDF} Collaboration, R.~D. Ball et~al., {\it {Parton distributions for
  the LHC Run II}},  {\em JHEP} {\bf 1504} (2015) 040,
  [\href{http://arxiv.org/abs/1410.8849}{{\tt arXiv:1410.8849}}].

\bibitem{Buckley:2014ana}
A.~Buckley, J.~Ferrando, S.~Lloyd, K.~Nordström, B.~Page, et~al., {\it
  {LHAPDF6: parton density access in the LHC precision era}},  {\em
  Eur.Phys.J.} {\bf C75} (2015), no.~3 132,
  [\href{http://arxiv.org/abs/1412.7420}{{\tt arXiv:1412.7420}}].

\bibitem{Aad:2014aba}
{\bf ATLAS} Collaboration, G.~Aad et~al., {\it {Measurement of the Higgs boson
  mass from the $H\rightarrow \gamma\gamma$ and $H \rightarrow ZZ^{*}
  \rightarrow 4\ell$ channels with the ATLAS detector using 25 fb$^{-1}$ of
  $pp$ collision data}},  {\em Phys.Rev.} {\bf D90} (2014), no.~5 052004,
  [\href{http://arxiv.org/abs/1406.3827}{{\tt arXiv:1406.3827}}].

\bibitem{Khachatryan:2014jba}
{\bf CMS} Collaboration, V.~Khachatryan et~al., {\it {Precise determination of
  the mass of the Higgs boson and tests of compatibility of its couplings with
  the standard model predictions using proton collisions at 7 and 8 $\,\text
  {TeV}$}},  {\em Eur.Phys.J.} {\bf C75} (2015), no.~5 212,
  [\href{http://arxiv.org/abs/1412.8662}{{\tt arXiv:1412.8662}}].

\bibitem{Aad:2015zhl}
{\bf ATLAS, CMS} Collaboration, G.~Aad et~al., {\it {Combined Measurement of
  the Higgs Boson Mass in $pp$ Collisions at $\sqrt{s}=7$ and 8 TeV with the
  ATLAS and CMS Experiments}},  {\em Phys. Rev. Lett.} {\bf 114} (2015) 191803,
  [\href{http://arxiv.org/abs/1503.07589}{{\tt arXiv:1503.07589}}].

\bibitem{Sjostrand:2007gs}
T.~Sjostrand, S.~Mrenna, and P.~Z. Skands, {\it {A Brief Introduction to PYTHIA
  8.1}},  {\em Comput.Phys.Commun.} {\bf 178} (2008) 852--867,
  [\href{http://arxiv.org/abs/0710.3820}{{\tt arXiv:0710.3820}}].

\bibitem{Sjostrand:2014zea}
T.~Sjöstrand, S.~Ask, J.~R. Christiansen, R.~Corke, N.~Desai, et~al., {\it {An
  Introduction to PYTHIA 8.2}},  {\em Comput.Phys.Commun.} {\bf 191} (2015)
  159--177, [\href{http://arxiv.org/abs/1410.3012}{{\tt arXiv:1410.3012}}].

\bibitem{Skands:2014pea}
P.~Skands, S.~Carrazza, and J.~Rojo, {\it {Tuning PYTHIA 8.1: the Monash 2013
  Tune}},  {\em Eur.Phys.J.} {\bf C74} (2014), no.~8 3024,
  [\href{http://arxiv.org/abs/1404.5630}{{\tt arXiv:1404.5630}}].

\bibitem{Ball:2012cx}
R.~D. Ball, V.~Bertone, S.~Carrazza, C.~S. Deans, L.~Del~Debbio, et~al., {\it
  {Parton distributions with LHC data}},  {\em Nucl.Phys.} {\bf B867} (2013)
  244--289, [\href{http://arxiv.org/abs/1207.1303}{{\tt arXiv:1207.1303}}].

\bibitem{Ball:2013hta}
{\bf NNPDF} Collaboration, R.~D. Ball, V.~Bertone, S.~Carrazza, L.~Del~Debbio,
  S.~Forte, A.~Guffanti, N.~P. Hartland, and J.~Rojo, {\it {Parton
  distributions with QED corrections}},  {\em Nucl. Phys.} {\bf B877} (2013)
  290--320, [\href{http://arxiv.org/abs/1308.0598}{{\tt arXiv:1308.0598}}].

\bibitem{Gleisberg:2008ta}
T.~Gleisberg et~al., {\it {Event generation with SHERPA 1.1}},  {\em JHEP} {\bf
  02} (2009) 007, [\href{http://arxiv.org/abs/0811.4622}{{\tt
  arXiv:0811.4622}}].

\bibitem{Bern:2011ep}
Z.~Bern, G.~Diana, L.~Dixon, F.~Febres~Cordero, S.~Hoeche, et~al., {\it
  {Four-Jet Production at the Large Hadron Collider at Next-to-Leading Order in
  QCD}},  {\em Phys.Rev.Lett.} {\bf 109} (2012) 042001,
  [\href{http://arxiv.org/abs/1112.3940}{{\tt arXiv:1112.3940}}].

\bibitem{Czakon:2013goa}
M.~Czakon, P.~Fiedler, and A.~Mitov, {\it {Total Top-Quark Pair-Production
  Cross Section at Hadron Colliders Through $O(α\frac{4}{S})$}},  {\em Phys.
  Rev. Lett.} {\bf 110} (2013) 252004,
  [\href{http://arxiv.org/abs/1303.6254}{{\tt arXiv:1303.6254}}].

\bibitem{Aad:2012gxa}
{\bf ATLAS} Collaboration, G.~Aad et~al., {\it {Search for the Standard Model
  Higgs boson produced in association with a vector boson and decaying to a
  $b$-quark pair with the ATLAS detector}},  {\em Phys.Lett.} {\bf B718} (2012)
  369--390, [\href{http://arxiv.org/abs/1207.0210}{{\tt arXiv:1207.0210}}].

\bibitem{Chatrchyan:2013zna}
{\bf CMS} Collaboration, S.~Chatrchyan et~al., {\it {Search for the standard
  model Higgs boson produced in association with a W or a Z boson and decaying
  to bottom quarks}},  {\em Phys.Rev.} {\bf D89} (2014), no.~1 012003,
  [\href{http://arxiv.org/abs/1310.3687}{{\tt arXiv:1310.3687}}].

\bibitem{Aad:2014xzb}
{\bf ATLAS} Collaboration, G.~Aad et~al., {\it {Search for the $b\bar{b}$ decay
  of the Standard Model Higgs boson in associated $(W/Z)H$ production with the
  ATLAS detector}},  {\em JHEP} {\bf 01} (2015) 069,
  [\href{http://arxiv.org/abs/1409.6212}{{\tt arXiv:1409.6212}}].

\bibitem{Cacciari:2011ma}
M.~Cacciari, G.~P. Salam, and G.~Soyez, {\it {FastJet User Manual}},  {\em
  Eur.Phys.J.} {\bf C72} (2012) 1896,
  [\href{http://arxiv.org/abs/1111.6097}{{\tt arXiv:1111.6097}}].

\bibitem{Cacciari:2005hq}
M.~Cacciari and G.~P. Salam, {\it {Dispelling the $N^{3}$ myth for the $k_t$
  jet-finder}},  {\em Phys. Lett.} {\bf B641} (2006) 57--61,
  [\href{http://arxiv.org/abs/hep-ph/0512210}{{\tt hep-ph/0512210}}].

\bibitem{Cacciari:2008gp}
M.~Cacciari, G.~P. Salam, and G.~Soyez, {\it {The Anti-k(t) jet clustering
  algorithm}},  {\em JHEP} {\bf 0804} (2008) 063,
  [\href{http://arxiv.org/abs/0802.1189}{{\tt arXiv:0802.1189}}].

\bibitem{Aad:2014bia}
{\bf ATLAS} Collaboration, G.~Aad et~al., {\it {Jet energy measurement and its
  systematic uncertainty in proton-proton collisions at $\sqrt{s}=7$ TeV with
  the ATLAS detector}},  {\em Eur. Phys. J.} {\bf C75} (2015), no.~1 17,
  [\href{http://arxiv.org/abs/1406.0076}{{\tt arXiv:1406.0076}}].

\bibitem{ATLAS:2012kla}
{\bf ATLAS} Collaboration, G.~Aad et~al., {\it {Performance of large-R jets and
  jet substructure reconstruction with the ATLAS detector}},
  \href{http://arxiv.org/abs/ATLAS-CONF-2012-065}{{\tt ATLAS-CONF-2012-065}}.

\bibitem{Butterworth:2008iy}
J.~M. Butterworth, A.~R. Davison, M.~Rubin, and G.~P. Salam, {\it {Jet
  substructure as a new Higgs search channel at the LHC}},  {\em
  Phys.Rev.Lett.} {\bf 100} (2008) 242001,
  [\href{http://arxiv.org/abs/0802.2470}{{\tt arXiv:0802.2470}}].

\bibitem{Dokshitzer:1997in}
Y.~L. Dokshitzer, G.~Leder, S.~Moretti, and B.~Webber, {\it {Better jet
  clustering algorithms}},  {\em JHEP} {\bf 9708} (1997) 001,
  [\href{http://arxiv.org/abs/hep-ph/9707323}{{\tt hep-ph/9707323}}].

\bibitem{Wobisch:1998wt}
M.~Wobisch and T.~Wengler, {\it {Hadronization corrections to jet
  cross-sections in deep inelastic scattering}},
  \href{http://arxiv.org/abs/hep-ph/9907280}{{\tt hep-ph/9907280}}.

\bibitem{Salam:2009jx}
G.~P. Salam, {\it {Towards Jetography}},  {\em Eur.Phys.J.} {\bf C67} (2010)
  637--686, [\href{http://arxiv.org/abs/0906.1833}{{\tt arXiv:0906.1833}}].

\bibitem{Aad:2013gja}
{\bf ATLAS} Collaboration, G.~Aad et~al., {\it {Performance of jet substructure
  techniques for large-$R$ jets in proton-proton collisions at $\sqrt{s}$ = 7
  TeV using the ATLAS detector}},  {\em JHEP} {\bf 1309} (2013) 076,
  [\href{http://arxiv.org/abs/1306.4945}{{\tt arXiv:1306.4945}}].

\bibitem{Butterworth:2002tt}
J.~Butterworth, B.~Cox, and J.~R. Forshaw, {\it {$W W$ scattering at the CERN
  LHC}},  {\em Phys.Rev.} {\bf D65} (2002) 096014,
  [\href{http://arxiv.org/abs/hep-ph/0201098}{{\tt hep-ph/0201098}}].

\bibitem{Ellis:1993tq}
S.~D. Ellis and D.~E. Soper, {\it {Successive combination jet algorithm for
  hadron collisions}},  {\em Phys. Rev.} {\bf D48} (1993) 3160--3166,
  [\href{http://arxiv.org/abs/hep-ph/9305266}{{\tt hep-ph/9305266}}].

\bibitem{Thaler:2010tr}
J.~Thaler and K.~Van~Tilburg, {\it {Identifying Boosted Objects with
  N-subjettiness}},  {\em JHEP} {\bf 1103} (2011) 015,
  [\href{http://arxiv.org/abs/1011.2268}{{\tt arXiv:1011.2268}}].

\bibitem{Thaler:2011gf}
J.~Thaler and K.~Van~Tilburg, {\it {Maximizing Boosted Top Identification by
  Minimizing N-subjettiness}},  {\em JHEP} {\bf 1202} (2012) 093,
  [\href{http://arxiv.org/abs/1108.2701}{{\tt arXiv:1108.2701}}].

\bibitem{Catani:1993hr}
S.~Catani, Y.~L. Dokshitzer, M.~H. Seymour, and B.~R. Webber, {\it
  {Longitudinally invariant $K\_t$ clustering algorithms for hadron hadron
  collisions}},  {\em Nucl. Phys.} {\bf B406} (1993) 187--224.

\bibitem{Larkoski:2013eya}
A.~J. Larkoski, G.~P. Salam, and J.~Thaler, {\it {Energy Correlation Functions
  for Jet Substructure}},  {\em JHEP} {\bf 1306} (2013) 108,
  [\href{http://arxiv.org/abs/1305.0007}{{\tt arXiv:1305.0007}}].

\bibitem{Larkoski:2014gra}
A.~J. Larkoski, I.~Moult, and D.~Neill, {\it {Power Counting to Better Jet
  Observables}},  {\em JHEP} {\bf 1412} (2014) 009,
  [\href{http://arxiv.org/abs/1409.6298}{{\tt arXiv:1409.6298}}].

\bibitem{Aad:2015ydr}
{\bf ATLAS} Collaboration, G.~Aad et~al., {\it {Performance of $b$-Jet
  Identification in the ATLAS Experiment}},
  \href{http://arxiv.org/abs/1512.01094}{{\tt arXiv:1512.01094}}.

\bibitem{Khachatryan:2011wq}
{\bf CMS} Collaboration, V.~Khachatryan et~al., {\it {Measurement of $B\bar{B}$
  Angular Correlations based on Secondary Vertex Reconstruction at $\sqrt{s}=7$
  TeV}},  {\em JHEP} {\bf 1103} (2011) 136,
  [\href{http://arxiv.org/abs/1102.3194}{{\tt arXiv:1102.3194}}].

\bibitem{Chatrchyan:2012jua}
{\bf CMS} Collaboration, S.~Chatrchyan et~al., {\it {Identification of b-quark
  jets with the CMS experiment}},  {\em JINST} {\bf 8} (2013) P04013,
  [\href{http://arxiv.org/abs/1211.4462}{{\tt arXiv:1211.4462}}].

\bibitem{Cacciari:2007fd}
M.~Cacciari and G.~P. Salam, {\it {Pileup subtraction using jet areas}},  {\em
  Phys. Lett.} {\bf B659} (2008) 119--126,
  [\href{http://arxiv.org/abs/0707.1378}{{\tt arXiv:0707.1378}}].

\bibitem{ATLAS-CONF-2014-004}
{\bf ATLAS} Collaboration, G.~Aad et~al., {\it {Calibration of $b$-tagging
  using dileptonic top pair events in a combinatorial likelihood approach with
  the ATLAS experiment}},  \href{http://arxiv.org/abs/ATLAS-CONF-2014-004}{{\tt
  ATLAS-CONF-2014-004}}.

\bibitem{Cacciari:2009dp}
M.~Cacciari, G.~P. Salam, and S.~Sapeta, {\it {On the characterisation of the
  underlying event}},  {\em JHEP} {\bf 1004} (2010) 065,
  [\href{http://arxiv.org/abs/0912.4926}{{\tt arXiv:0912.4926}}].

\bibitem{TheATLAScollaboration:2013pia}
{\bf ATLAS} Collaboration, G.~Aad et~al., {\it {Pile-up subtraction and
  suppression for jets in ATLAS}},
  \href{http://arxiv.org/abs/ATLAS-CONF-2013-083}{{\tt ATLAS-CONF-2013-083}}.

\bibitem{Krohn:2009th}
D.~Krohn, J.~Thaler, and L.-T. Wang, {\it {Jet Trimming}},  {\em JHEP} {\bf 02}
  (2010) 084, [\href{http://arxiv.org/abs/0912.1342}{{\tt arXiv:0912.1342}}].

\bibitem{Krohn:2013lba}
D.~Krohn, M.~D. Schwartz, M.~Low, and L.-T. Wang, {\it {Jet Cleansing: Pileup
  Removal at High Luminosity}},  {\em Phys.Rev.} {\bf D90} (2014), no.~6
  065020, [\href{http://arxiv.org/abs/1309.4777}{{\tt arXiv:1309.4777}}].

\bibitem{Cacciari:2008gd}
M.~Cacciari, J.~Rojo, G.~P. Salam, and G.~Soyez, {\it {Quantifying the
  performance of jet definitions for kinematic reconstruction at the LHC}},
  {\em JHEP} {\bf 12} (2008) 032, [\href{http://arxiv.org/abs/0810.1304}{{\tt
  arXiv:0810.1304}}].

\bibitem{Ellis:2009me}
S.~D. Ellis, C.~K. Vermilion, and J.~R. Walsh, {\it {Recombination Algorithms
  and Jet Substructure: Pruning as a Tool for Heavy Particle Searches}},  {\em
  Phys. Rev.} {\bf D81} (2010) 094023,
  [\href{http://arxiv.org/abs/0912.0033}{{\tt arXiv:0912.0033}}].

\bibitem{Bertolini:2014bba}
D.~Bertolini, P.~Harris, M.~Low, and N.~Tran, {\it {Pileup Per Particle
  Identification}},  {\em JHEP} {\bf 10} (2014) 59,
  [\href{http://arxiv.org/abs/1407.6013}{{\tt arXiv:1407.6013}}].

\bibitem{Cacciari:2014gra}
M.~Cacciari, G.~P. Salam, and G.~Soyez, {\it {SoftKiller, a particle-level
  pileup removal method}},  {\em Eur. Phys. J.} {\bf C75} (2015), no.~2 59,
  [\href{http://arxiv.org/abs/1407.0408}{{\tt arXiv:1407.0408}}].

\bibitem{Cacciari:2014jta}
M.~Cacciari, G.~P. Salam, and G.~Soyez, {\it {Use of charged-track information
  to subtract neutral pileup}},  {\em Phys. Rev.} {\bf D92} (2015), no.~1
  014003, [\href{http://arxiv.org/abs/1404.7353}{{\tt arXiv:1404.7353}}].

\bibitem{Berta:2014eza}
P.~Berta, M.~Spousta, D.~W. Miller, and R.~Leitner, {\it {Particle-level pileup
  subtraction for jets and jet shapes}},  {\em JHEP} {\bf 06} (2014) 092,
  [\href{http://arxiv.org/abs/1403.3108}{{\tt arXiv:1403.3108}}].

\bibitem{Larkoski:2014wba}
A.~J. Larkoski, S.~Marzani, G.~Soyez, and J.~Thaler, {\it {Soft Drop}},  {\em
  JHEP} {\bf 05} (2014) 146, [\href{http://arxiv.org/abs/1402.2657}{{\tt
  arXiv:1402.2657}}].

\bibitem{Cacciari:2010te}
M.~Cacciari, J.~Rojo, G.~P. Salam, and G.~Soyez, {\it {Jet Reconstruction in
  Heavy Ion Collisions}},  {\em Eur. Phys. J.} {\bf C71} (2011) 1539,
  [\href{http://arxiv.org/abs/1010.1759}{{\tt arXiv:1010.1759}}].

\bibitem{Soper:2011cr}
D.~E. Soper and M.~Spannowsky, {\it {Finding physics signals with shower
  deconstruction}},  {\em Phys.Rev.} {\bf D84} (2011) 074002,
  [\href{http://arxiv.org/abs/1102.3480}{{\tt arXiv:1102.3480}}].

\bibitem{Soper:2012pb}
D.~E. Soper and M.~Spannowsky, {\it {Finding top quarks with shower
  deconstruction}},  {\em Phys.Rev.} {\bf D87} (2013) 054012,
  [\href{http://arxiv.org/abs/1211.3140}{{\tt arXiv:1211.3140}}].

\bibitem{Baldi:2014pta}
P.~Baldi, P.~Sadowski, and D.~Whiteson, {\it {Enhanced Higgs Boson to
  $\tau^+\tau^-$ Search with Deep Learning}},  {\em Phys.Rev.Lett.} {\bf 114}
  (2015), no.~11 111801, [\href{http://arxiv.org/abs/1410.3469}{{\tt
  arXiv:1410.3469}}].

\bibitem{Aaltonen:2012qt}
  {\bf CDF} and {\bf D0}
  Collaboration, T.~Aaltonen et~al., {\it {Evidence for a particle
  produced in association with weak bosons and decaying to a bottom-antibottom
  quark pair in Higgs boson searches at the Tevatron}},  {\em Phys. Rev. Lett.}
  {\bf 109} (2012) 071804, [\href{http://arxiv.org/abs/1207.6436}{{\tt
        arXiv:1207.6436}}].

\bibitem{Kang:2015uoc} 
  Z.~Kang, P.~Ko and J.~Li,
  {\it New Avenues to Heavy Right-handed Neutrinos with Pair Production at Hadronic Colliders},
   \href{http://arxiv.org/abs/1512.08373}{{\tt
       arXiv:1512.08373}}.

\bibitem{DelDebbio:2004qj}
{\bf The NNPDF} Collaboration, L.~Del~Debbio, S.~Forte, J.~I. Latorre,
  A.~Piccione, and J.~Rojo, {\it Unbiased determination of the proton structure
  function f2(p) with faithful uncertainty estimation},  {\em JHEP} {\bf 03}
  (2005) 080, [\href{http://arxiv.org/abs/hep-ph/0501067}{{\tt
  hep-ph/0501067}}].

\bibitem{Ball:2008by}
{\bf The NNPDF} Collaboration, R.~D. Ball et~al., {\it {A determination of
  parton distributions with faithful uncertainty estimation}},  {\em Nucl.
  Phys.} {\bf B809} (2009) 1--63, [\href{http://arxiv.org/abs/0808.1231}{{\tt
  arXiv:0808.1231}}].

\bibitem{Ball:2011mu}
{\bf The NNPDF} Collaboration, R.~D. Ball et~al., {\it {Impact of Heavy Quark
  Masses on Parton Distributions and LHC Phenomenology}},  {\em Nucl. Phys.}
  {\bf B849} (2011) 296--363, [\href{http://arxiv.org/abs/1101.1300}{{\tt
  arXiv:1101.1300}}].

\bibitem{Ball:2010de}
{\bf {The NNPDF}} Collaboration, R.~D. Ball et~al., {\it {A first unbiased
  global NLO determination of parton distributions and their uncertainties}},
  {\em Nucl. Phys.} {\bf B838} (2010) 136--206,
  [\href{http://arxiv.org/abs/1002.4407}{{\tt arXiv:1002.4407}}].

\bibitem{quevedo}
B.~C. Allanach, D.~Grellscheid, and F.~Quevedo, {\it Genetic algorithms and
  experimental discrimination of susy models},  {\em JHEP} {\bf 07} (2004) 069,
  [\href{http://arxiv.org/abs/hep-ph/0406277}{{\tt hep-ph/0406277}}].

\bibitem{tau}
J.~Rojo and J.~I. Latorre, {\it Neural network parametrization of spectral
  functions from hadronic tau decays and determination of qcd vacuum
  condensates},  {\em JHEP} {\bf 01} (2004) 055,
  [\href{http://arxiv.org/abs/hep-ph/0401047}{{\tt hep-ph/0401047}}].

\bibitem{Abel:2014xta}
S.~Abel and J.~Rizos, {\it {Genetic Algorithms and the Search for Viable String
  Vacua}},  {\em JHEP} {\bf 08} (2014) 010,
  [\href{http://arxiv.org/abs/1404.7359}{{\tt arXiv:1404.7359}}].

\bibitem{Nesseris:2012tt}
S.~Nesseris and J.~Garcia-Bellido, {\it {A new perspective on Dark Energy
  modeling via Genetic Algorithms}},  {\em JCAP} {\bf 1211} (2012) 033,
  [\href{http://arxiv.org/abs/1205.0364}{{\tt arXiv:1205.0364}}].

\end{thebibliography}\endgroup

\end{document}